I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-06-08 paper_io_sim.tex
Comments up to Algo 1

	\section{Method}
		In this section, we discuss our approach to model page cache and
		the cache eviction strategy implemented in Linux. We also detail
		the design the simulators in order to generalize page cache as well
		as the implementation in Python and SimGrid framework. Finally, we
		describe some experimental scenarios to compare the results of
		Python simulator, baseline SimGrid simulator and the results from
		the real pipelines.

@@

			When modeling the I/O time, we consider the key factors of the
			I/O mechanism having the biggest impacts on data read/write.
			The first factor taken into account is the use of the page
			cache in Linux kernel. The second factor  is the cache eviction, 
			data flushing and cache eviction mechanisms implemented in the 
			Linux kernel, as they manage and change the status of the memory 
			including cache size, amount of available memory and 
			dirty data, based on which the kernel controls I/O activities. Therefore, 
			the idea of our simulator is to simulate the memory, the Linux 
			page cache with its LRU lists, storage device and imitate the data flushing 
			and cache eviction algorithm implemented in the kernel.

			At the level of detail in our simulator, the simulation models of memory 
			and storage devices are straightforward when they can be modeled with 
			predefined capacity and bandwidth. To simulate the cache eviction mechanism 
			implemented in Linux, our simulator also keeps tracks of page cache 
			LRU lists as in the kernel. However, instead of dealing with lists pages, we 
			consider a sequence of continuously accessed pages of a file as a single 
			unit, a block of data. A data block keeps the information about file name, 
			block size, last access and a dirty flag describing whether the data is 
			dirty or not.

@@

			Having these simulation components, we can generalize I/O activities in
			our simulator. When the kernel reads and writes data, it simply 
			reads/writes data from/to disk or cache with predefined disk bandwidth 
			or memory bandwidth. The data blocks are added or updated in page cache 
			LRU lists accordingly. In flushing and periodical flushing, the kernel 
			traverses through the inactive and active lists to reduce the amount of 
			dirty data, while cache eviction algorithm deletes blocks from the 
			inactive list. The page cache LRU lists are updated after every I/O action.

@@

			We encapsulate our simulation components with kernel-like functionalities 
			into a layer between user simulated application and storage simulation 
			component (Figure~\ref{fig:interaction}). Our new layer includes two main
			classes, the \textbf{\textit{IOController}}, which models the kernel 
			I/O controlling logic, and the \textbf{\textit{MemoryManager}}, which 
			manages page cache LRU lists and stores all necessary information about 
			memory including capacity, amount of cache, amount of memory used 
			by applications. Because every node has its own memory, storage devices 
			and system configurations, there is one IOController object and one 
			MemoryManager object created for each node. The IOController class 
			keeps system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), 
			implements necessary I/O control functionalities such as read, write,
			flushing, and cache evicting.

			With this layer, user simulated applications send requests to an 
			IOController of the node where requested data is stored, instead of 
			sending read/write requests directly to storage model as in simulators 
			with the current SimGrid version. Based on the memory status in 
			MemoryManager, the IOController orchestrates cache read/write, 
			disk read/write, flushing and cache eviction to fulfill application 
			requests. The time for each action is estimated and the total time of 
			a request is returned to user applications.

			\begin{algorithm}\caption{Read}\label{alg:read}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{fn}{file name}
        				\Desc{fs}{file size}
						\Desc{rt}{current simulated run time}
						\Desc{mm}{MemoryManager object}
						\Desc{$r_m$}{memory read bandwidth}
						\Desc{$r_d$}{disk read bandwidth}
   					\EndInput
   					\Output
						\Desc{rt}{simulated run time after the file is read}
   					\EndOutput
					\State mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.available\_memory()) 
					\State mem\_avail = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.dirty\_data())
					\State rt = rt +  mm.flush(mem\_required - 
					mem\_avail)	
					\State free\_mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.free\_memory())
					\State mm.evict(free\_mem\_required) 
					\If {mm.cached\_data(fn) $>$ 0}
    					\State mm.cache\_read(fn) 
						\If {mm.cached\_data(fn) == fs}
    						\State time = mm.cache\_data(fn) / $r_m$
    						\State IOController.periodic\_flush(time) 
							\State rt = rt + time
						\EndIf
					\EndIf
					\If {mm.cached\_data(fn) $<$ fs}
						\State rt = rt + IOController.flush(mm.dirty\_data())
						\State mm.disk\_read(fs - mm.cache\_data(fn), fn)
    					\State rt = rt + (fs - mm.cache\_data(fn)) / $r_d$
					\EndIf

@@

			The algorithm to generalize file read is described in 
			\textit{Algorithm~\ref{alg:read}}. In Linux systems, data can be read in 
			two ways, read from cache with memory bandwidth and read from disk 
			with disk bandwidth. Because total read time can be seen as a 
			linear function of the read time in each of these two modes, 
			we can generalize reading process as two separate and sequential phases, 
			one is cache read, and the other is disk read. When a user application 
			requests a file read, it calls the read method of the IOController. Then, 
			IOController calculates how much memory is required to read the file 
			based on the amount of available memory, file size, and the amount of 
			file data cached (line 10). If there is not enough available memory, 
			dirty data is flushed to disk to make more memory available by calling flush
			function with the amount of memory needed as the parameter (line 11-12).
			IOController then evicts old blocks from the inactive list to
			accommodate the requested file and flushing time is simulated (line 13-14). 
			After that, IOController starts estimating file read time by first checking 
			if a some file data is cached (line 15). If there is cached data, 
			it is read from cache and the blocks containing this data are updated in 
			the page cache LRU lists with the last access time (line 16). If all file data 
			is cached, time to read from cache is added to total run time, and 
			periodical flushing is triggered to flush dirty data during the cache read 
			(line 18-19). If only a part of the file is in cache, disk read is required, 
			this does not allow flushing and makes the cache read time neglectable. 
			At the beginning of a disk read, dirty data is flushed, then data is read 
			and added to the inactive list, disk read time is estimated with disk read 
			bandwidth and added to total run time (line 23-27).

			\textit{Algorithm~\ref{alg:write}} describes our
			generalization of file write. Similar to the file read process,
			a file can also be written with two different bandwidths,
			memory bandwidth and disk bandwidth, and thus file write can be
			generalized with two modes. File can be written with memory
			bandwidth before the page cache is saturated, or the amount of
			dirty data reaches dirty\_ratio. Firstly, if dirty\_ratio is not reached 
			the amount of data can be written with memory bandwidth is
			calculated with the formula:
			\begin{equation}
				M = max(0, \frac{(d*C - D)*w_m}{w_m - w_d})\label{equa:freeamt}
			\end{equation}

>>>>

	\section{Method}
		In this section, we present our approach to model page cache and
		the cache eviction strategy implemented in Linux. We also detail
		the design the simulators in order to generalize page cache as well
		as the implementation in Python and SimGrid framework. Finally, we
		describe some experimental scenarios to compare the results of
		Python simulator, baseline SimGrid simulator and the results from
		the real pipelines.

@@

			When modeling the I/O time, we consider the key factors of the
			I/O mechanism having the biggest impacts on data read/write.
			The first factor taken into account is the use of the page
			cache in Linux kernel. The second factor  is the cache eviction, 
			data flushing and cache eviction mechanisms implemented in the 
			Linux kernel, as they manage and change the status of the memory 
			including cache size, amount of available memory and 
			dirty data, based on which the kernel controls I/O activities. Therefore, 
			the idea of our simulator is to simulate the memory, the Linux 
			page cache with its LRU lists, storage device and imitate the data flushing 
			and cache eviction algorithm implemented in the kernel. \tristan{This paragraph is convoluted, it needs to be rewritten with a straightforward logic.}

			At the level of detail in our simulator, the simulation models of memory 
			and storage devices are straightforward when they can be modeled with 
			predefined capacity and bandwidth \tristan{Do you mean ``We
			model memory as a storage device, with a capacity and a
			bandwidth''? If that's the case, write this sentence like this,
			it's more direct and less ambiguous. You should also explain
			how you calculate read/write time from the bandwidth. And does
			your model distinguish read and write bandwidths? I think the
			memory model is one paragraph, not just one sentence in a
			paragraph that then talks about other things}. To simulate the
			cache eviction mechanism implemented in Linux, our simulator
			also keeps tracks of page cache LRU lists as in the kernel.
			However, instead of dealing with lists pages \tristan{as done by the kernel?}, we 
			consider a sequence of continuously accessed pages of a file as a single 
			unit, a block of data \tristan{so is a data block a set of pages? The definition of a data block is unclear.}. A data block keeps the information about file name, 
			block size, last access and a dirty flag describing whether the data is 
			dirty or not.

@@

			Having these simulation components, we can generalize I/O activities in
			our simulator. When the kernel reads and writes data, it simply 
			reads/writes data \tristan{do you mean data blocks?} from/to disk or cache with predefined disk bandwidth 
			or memory bandwidth. The data blocks are added or updated in page cache 
			LRU lists accordingly. In flushing and periodical flushing, the kernel 
			traverses through the inactive and active lists to reduce the amount of 
			dirty data, while cache eviction algorithm deletes blocks from the 
			inactive list. The page cache LRU lists are updated after every I/O action.

@@

			We encapsulate our simulation components with kernel-like functionalities 
			into a layer between the simulated application and the simulated storage (Figure~\ref{fig:interaction}). Our new layer includes two main
			classes, the \emph{IOController}, which models the kernel 
			I/O controlling logic, and the \emph{MemoryManager}, which 
			manages page cache LRU lists and stores all necessary information about 
			memory including total capacity, current cache size, and current amount of memory used 
			by applications \tristan{is it exhaustive?}. Because every node has its own memory, storage devices 
			and system configurations, one IOController object and one 
			MemoryManager object are created for each node. The IOController class 
			stores system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), 
			implements necessary I/O control functionalities such as read, write,
			flushing, and cache eviction.

			Simulated applications send requests to an IOController of the
			node where the requested data is stored, instead of sending
			read/write requests directly to storage model as in the current
			SimGrid version. Based on the memory status in 
			MemoryManager, the IOController orchestrates cache read/write, 
			disk read/write, flushing and cache eviction to fulfill application 
			requests. The time for each action is estimated and the total time of 
			a request is returned to application.

			\begin{algorithm}\caption{Read \tristan{Find a more descriptive title. Lines 10, 11 and 13 should fit on a single line. Suggestion: replace cached\_data with cached, available\_memory with avail.
				Add comments in the code. Line 23: should it be in an else clause? } }\label{alg:read}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{fn}{file name}
        				\Desc{fs}{file size}
						\Desc{rt}{current simulated run time}
						\Desc{mm}{MemoryManager object}
						\Desc{$r_m$}{memory read bandwidth}
						\Desc{$r_d$}{disk read bandwidth}
   					\EndInput
   					\Output
						\Desc{rt}{simulated run time after the file is read}
   					\EndOutput
					\State mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.available\_memory()) 
					\State mem\_avail = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.dirty\_data())
					\State rt = rt +  mm.flush(mem\_required - 
					mem\_avail)	
					\State free\_mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.free\_memory())
					\State mm.evict(free\_mem\_required) 
					\If {mm.cached\_data(fn) $>$ 0}
    					\State mm.cache\_read(fn) 
						\If {mm.cached\_data(fn) == fs}
    						\State time = mm.cache\_data(fn) / $r_m$
    						\State IOController.periodic\_flush(time) 
							\State rt = rt + time
						\EndIf
					\EndIf
					\If {mm.cached\_data(fn) $<$ fs}
						\State rt = rt + IOController.flush(mm.dirty\_data())
						\State mm.disk\_read(fs - mm.cache\_data(fn), fn)
    					\State rt = rt + (fs - mm.cache\_data(fn)) / $r_d$
					\EndIf

					\State \Return rt

@@

			The algorithm to generalize \tristan{``generalize'' isn't a great word here, I'd simply say ``simulate''} file reads is described in 
			Algorithm~\ref{alg:read}. In Linux systems, data can be read in 
			two ways, read from cache with memory bandwidth and read from disk 
			with disk bandwidth. Because total read time can be seen as a 
			linear function of the read time in each of these two modes, 
			we can generalize reading process as two separate and sequential phases, 
			one is cache read, and the other is disk read. When a user application 
			requests a file read, it calls the read method of the IOController. Then, 
			IOController calculates how much memory is required to read the file 
			based on the amount of available memory, file size, and the amount of 
			file data cached (line 10). If there is not enough available memory, 
			dirty data is flushed to disk to make more memory available by calling flush
			function with the amount of memory needed as the parameter (line 11-12).
			IOController then evicts old blocks \tristan{I don't see blocks being handled in this algo. It adds to the confusion related to the 
			lack of a proper definition of blocks.} from the inactive list to
			accommodate the requested file and flushing time is simulated (line 13-14). 
			After that, IOController starts estimating file read time by first checking 
			if a some file data is cached (line 15). If there is cached data, 
			it is read from cache and the blocks containing this data are updated in 
			the page cache LRU lists with the last access time (line 16). If all file data 
			is cached, time to read from cache is added to total run time, and 
			periodical flushing is triggered to flush dirty data during the cache read 
			(line 18-19). If only a part of the file is in cache, disk read is required, 
			this does not allow flushing and makes the cache read time neglectable. 
			At the beginning of a disk read, dirty data is flushed, then data is read 
			and added to the inactive list, disk read time is estimated with disk read 
			bandwidth and added to total run time (line 23-27).

			Algorithm~\ref{alg:write} describes our
			generalization of file write. Similar to the file read process,
			a file can also be written with two different bandwidths,
			memory bandwidth and disk bandwidth, and thus file write can be
			generalized with two modes. File can be written with memory
			bandwidth before the page cache is saturated, or the amount of
			dirty data reaches dirty\_ratio. Firstly, if dirty\_ratio is not reached 
			the amount of data can be written with memory bandwidth is
			calculated with the formula:
			\begin{equation}
				M = max(0, \frac{(d*C - D)*w_m}{w_m - w_d})\label{equa:freeamt}
			\end{equation}
